Untitled

### 2020-10-20 ms/manuscript.Rmd
Read in climate data

---
title: ""
bibliography: references.bib
mainfont: Roboto-Regular.ttf
fontsize: 12pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
---

<!-- to move to the next line without starting a new paragraph, 
add two spaces after last word, e.g., Atsushi Ebihara__ -->
**Corresponding author**  
Joel H. Nitta  
Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan  
Phone: +81-  
Fax: +81-  
joelnitta@gmail.com

**Membership of the Botanical Society of Japan**  
YES (Joel H. Nitta)

**Subject area**  
(1)	Taxonomy/Phylogenetics/Evolutionary Biology

**Number of tables, black-and-white figures, and color figures.**  
No tables / 2 black-and-white figure / 1 color figure

<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
for inserting page breaks when rendering to MS word -->
#####
\newpage

@@

ORCID ID numbers  
- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}  
- N. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}  
- Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}

### Abstract
Desiccation tolerance was a key trait allowing plants to colonize land.
However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits. 
Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
Epiphytes were more tolerant of desiccation than terrestrial species.
Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.

@@

#####
\newpage

@@

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often <1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

@@

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, [@Shreve1911] demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, @Johnson2000 found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
@Parra2009 and @Saldana2013 investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].

Despite this considerable interest in filmy fern ecophysiology, with the sole exception of @Johnson2000, no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Proctor2003; @Parra2015; @Flores-Bavestrello2016].
Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes  [@Benzing1990; @WatkinsJr.2009].
In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
3.) For filmy ferns with gametophytes that occur over a wider range than sporophytes, the gametophyte has higher DT and photosynthetic optima than the sporophyte.

@@

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca. 15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea [@Murdock2003; @Ranker2005; @Nitta2011b; @Florence], including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
We follow the taxonomic system of @Ebihara2006.

@@

<!-- 
### Microclimate

We characterized the microenvironment experienced by filmy ferns by logging relative humidity (RH) and temperature every 15 min from July 7, 2013 to July 5, 2014 at 16 of the sampling sites (Fig. 2).
To do this we used Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA) with external sensors outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation.
To compare between terrestrial and epiphytic habitats, dataloggers were placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at ca. 10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
We excluded any days during which one or more dataloggers failed from the final dataset, and some sites are missing data completely for one or both dataloggers.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013).
We used the previous year's data for this site instead.
The final dataset included 201 days of data from 15 sites (26 dataloggers).
We tested for differences in climatic variables between epiphytic and terrestrial habitats using analysis of covariance (ANCOVA) with each climatic variable in turn as the response variable and elevation as the explanatory variable with habitat (terrestrial vs. epiphytic) as a cofactor.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval (Testo and Watkins, 2013).
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22--25 *◦*C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (--282 MPa), Mg(NO~3~)~2~ (--86 MPa), and NaCl (--38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at --86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0*.*5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield (Watkins et al., 2007).

@@

Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400--700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0*.*8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^--2^ ·s^--1^.
Response curves were fitted using the equation *y* = *A*(1 --*e^-kx^*), where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded (Proctor, 2012).
The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

@@

We used DNA barcoding to identify filmy fern gametophytes to species as described in Nitta et al. (2016).
Briefly, DNA extraction was performed using a modified CTAB protocol (Doyle and Doyle, 1987) or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following Nitta et al. (2016).
For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST (Altschul et al., 1997).
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching *\>*99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched *\>*99.5% with multiple reference sequences were not identified and excluded from further analysis.

@@

All statistical analyses were done in R v.3.2.1 (R Core Team, 2015).
Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).

We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of Nitta et al. (2016) with the extract.clade function of the ape package in R (Paradis et al., 2004).
This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies (Pryer et al., 2001; Dubuisson et al., 2003; Hennequin et al., 2006; Ebihara et al., 2007; Appendix S1, see Supplemental Data with the online version of this article).

We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* (Blomberg et al., 2003) and Pagel's λ (Pagel, 1999; Freckleton et al., 2002) as implemented with the phylosig function in the phytools R package (Revell, 2012).
Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
λ is a scaling parameter that ranges from zero to one: λ near zero indicates random distribution of trait values across the tree, and λ near one indicates evolution of traits along the phylogeny following BM.
For *K*, values near one indicate evolution of traits following BM; *K* *\>*1 indicates traits more conserved than expected under BM, and *K* *\<*1 indicates that traits have less phylogenetic signal than expected under BM.
We tested the significance of *K* by comparing the observed value against values from a null distribution of 1000 phylogenies with the traits randomly shuffled across the tips.
We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal, and standard methods otherwise.

We used general linear mixed models (GLMMs) to investigate the effect of growth habit, range size, and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at --86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs. non-epiphyte $$i.e., terrestrial or saxicolous$$) or range size (widespread vs. not widespread) and generation (sporophyte vs. gametophyte) and their interactions as fixed effects.
Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix only for response variables with significant phylogenetic signal (Hadfield, 2010).
We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
Comparisons of DT, ETR~max~ , and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.

To test for a correlation between environmental water potential and DT, we first derived a single minimum water potential value for each species by taking the average of the minimum daily water potential at all sites where that species was present.
We used minimum values because these may represent a climatic extreme that limits the occurrence of species.
We used data from the 'epiphytic' dataloggers mounted at 1.5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
Data from the Mt. Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data (see Results).
This may be due to its location; unlike other mountains on Moorea, Mt. Rotui is completely isolated with no connecting ridges to other peaks (Fig. 2).
Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.
We ran PGLS using the caper R package (Orme et al., 2013).

@@

-->

## References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->

<div id="refs"></div>

>>>>

---
title: ""
bibliography: references.bib
mainfont: Roboto-Regular.ttf
fontsize: 12pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none # Use pandoc default, otherwise csl can't be used for reference formatting
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
    pandoc_args: [ "--csl", "journal-of-plant-research.csl"]
    includes:
      in_header: journal-of-plant-research.sty
header-includes: 
  \usepackage{float}
  \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r setup}
library(jntools)
```

```{r load-data}
loadd(
  list = c(
    "climate"
  )
)
```

```{=html}
<!-- to move to the next line without starting a new paragraph, 
add two spaces after last word, e.g., Atsushi Ebihara__ -->
```
**Corresponding author**\
Joel H.
Nitta\
Department of Biological Sciences, Graduate School of Science, The University of Tokyo, 2-11-16 Yayoi, Bunkyo-ku, Tokyo, 113-0032, Japan\
Phone: +81-\
Fax: +81-\
[joelnitta\@gmail.com](mailto:joelnitta@gmail.com){.email}

**Membership of the Botanical Society of Japan**\
YES (Joel H. Nitta)

**Subject area**\
(1) Taxonomy/Phylogenetics/Evolutionary Biology

**Number of tables, black-and-white figures, and color figures.**\
No tables / 2 black-and-white figure / 1 color figure

```{=html}
<!-- See https://stackoverflow.com/questions/24672111/how-to-add-a-page-break-in-word-document-generated-by-rstudio-markdown/40474753
for inserting page breaks when rendering to MS word -->
```
#####

\newpage

@@

ORCID ID numbers\
- Joel H. Nitta: \textnormal{https://orcid.org/0000-0003-4719-7472}\
- N
. Michele Holbrook: \textnormal{https://orcid.org/0000-0003-3325-5395}\
- Charles C. Davis: \textnormal{https://orcid.org/0000-0001-8747-1101}

### Abstract

Desiccation tolerance was a key trait allowing plants to colonize land.
However, little is known about the transition from desiccation tolerant non-vascular plants to desiccation sensitive vascular ones.
Filmy ferns (Hymenophyllaceae) present a useful system to study how water-stress strategies differ between non-vascular and vascular stages within a single organism, as they have vascularized sporophytes and nonvascular gametophytes that are each capable of varying degrees of desiccation tolerance.
We surveyed sporophytes and gametophytes of 22 species of filmy ferns on Moorea (French Polynesia) and used chlorophyll fluorescence to measure desiccation tolerance and light responses.
We sequenced plastid *rbcL* and *trnHpsbA* to identify field-collected gametophytes, and conducted phylogenetically informed analyses to identify differences in physiology between life stages and growth habits.
Gametophytes had similar or less desiccation tolerance and lower photosynthetic optima than sporophytes.
Epiphytes were more tolerant of desiccation than terrestrial species.
Despite their lack of greater physiological tolerances, gametophytes of several species occurred over a wider elevational range than conspecific sporophytes.
Our results show that filmy fern gametophytes and sporophytes differ in their physiology and niche requirements, and point to the importance of microhabitat in shaping the evolution of water-use strategies in vascular plants.

@@

#####

\newpage

@@

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often \<1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013].
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

@@

The degree of DT in filmy fern sporophytes occupying different niches has been relatively well studied.
In a series of elegant experiments on Jamaican filmy ferns, [@Shreve1911] demonstrated that water-use strategies differ between species occupying contrasting habitats (i.e., terrestrial, low elevation epiphyte, and high elevation epiphyte).
Proctor [-@Proctor2003; -@Proctor2012] investigated water relations and the ability to recover from desiccation in sporophytes of several filmy ferns, and found a general correlation between high DT and adaptation for high light levels in widespread species vs. low DT and adaptation to low light levels in species occupying more sheltered habitats.
In the only study to our knowledge including in situ physiological observations of hymenophyllaceous gametophytes, [@Johnson2000] found that gametophytes of *Vandenboschia speciosa* (Willd.) G.
Kunkel (= *Trichomanes speciosum* Willd.) are adapted for extremely low light environments, and operate at lower light levels than conspecific sporophytes.
[@Parra2009] and [@Saldana2013] investigated the relationship between physiology (photosynthetic rates and water relations) and vertical stratification in several epiphytic filmy fern species (primarily genus *Hymenophyllum*) along tree trunks in Chile.
They found that species occurring at greater heights in the trees tended to have greater DT, higher photosynthetic capacity (A~max~), and lower rates of evapotranspiration.
A recent study using two species of *Hymenophyllum* demonstrated that DT in filmy ferns likely operates via a constitutive mechanism, which is more similar to that of bryophytes than other drought-tolerant vascular plants, thus allowing filmy ferns to adjust to dry conditions extremely rapidly [@GarcesCea2014].

Despite this considerable interest in filmy fern ecophysiology, with the sole exception of [@Johnson2000], no studies have compared physiology between gametophyte and sporophyte phases, and we are aware of none that have done so across multiple species.
Furthermore, to our knowledge, no physiological studies of filmy ferns have accounted for their phylogenetic relationships, and we lack an understanding of how physiological strategies have evolved in this clade.
In addition to DT, photosynthetic optima (adaptation to different light levels) are likely to be important in determining the niches of life stages and species of filmy ferns [@Proctor2003; @Parra2015; @Flores-Bavestrello2016].
Terrestrial and epiphytic species experience different levels of photosynthetic radiation and desiccation, and light intensity varies with height along the host plant in epiphytes [@Benzing1990; @WatkinsJr.2009].
In a recent study (Testo et al., unpublished), the combined effect of desiccation and light stress was found to have a greater impact on filmy fern physiological performance than either of these stressors alone, suggesting the need to integrate studies of DT with photosynthesis to gain a full picture of physiological adaptation in filmy ferns.
Here, we leverage a recently developed DNA barcoding system [@Nitta2017] to test several hypotheses related to DT and photosynthesis in a comparative phylogenetic framework including both filmy fern sporophytes and gametophytes: 1.) Epiphytic filmy ferns have higher DT and photosynthetic optima than non-epiphytic species.
2.) Amongst epiphytic filmy ferns, those occurring at high elevation have higher DT and photosynthetic optima than those at low elevation.
3.) Gametophytes with wider ranges have higher DT than those with with narrow ranges.

@@

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology (Fig. 2).
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located ca.
15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
At the time of the study, a total of 18 filmy fern species were known from Moorea [@Murdock2003; @Ranker2005; @Nitta2011b; @Florence], including multiple terrestrial, saxicolous, and epiphytic species.
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by similar numbers of species.
We follow the taxonomic system of [@Ebihara2006].

@@

```{r microclimate-stats}
climate_start <- climate %>% pull(date) %>% min()
climate_end <- climate %>% pull(date) %>% max()
climate_n_sites <- climate %>% pull(site) %>% n_distinct()
climate_n_days <- climate %>% pull(date) %>% n_distinct()
climate_n_loggers <- climate %>% count(site, habit) %>% nrow()
```

We used microclimate data collected by [@Nitta2020].
This dataset includes relative humidity (RH) and temperature measured every 15 min from `r climate_start` to `r climate_end` at the sampling sites with Hobo ProV2 dataloggers (Onset Computer Corporation, Bourne, Massachusetts, USA).
The dataloggers were outfitted with radiation shields (RS1 or RS3, Onset Computer Corporation, Bourne, Massachusetts, USA) to prevent direct contact with precipitation and solar radiation, and placed in pairs at each site (one 'epiphytic' datalogger at ca. 1.5 m on a tree, one 'terrestrial' datalogger at ca. 10 cm next to the tree).
During the study period, some dataloggers malfunctioned, presumably due to prolonged exposure to high humidity.
Any days during which one or more dataloggers failed are excluded from the dataset.
One terrestrial datalogger (Mt. Tohiea 393 m site) was lacking data for more than half of the survey period, but had data available for the same period from a preliminary survey conducted during the previous year (2012--2013), so the previous year's data for this site was used instead.
The dataset includes `r climate_n_days` days of data from `r climate_n_sites` sites (`r climate_n_loggers` dataloggers).

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8--16 individuals (one individual = single whole frond including a ca. 2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment chlorophyll fluorescence (F~v~/F~m~) was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Plants were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100% RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Hobo ProV2 dataloggers.
The desiccation chambers were kept in an air-conditioned room, and temperature inside the chambers ranged from 22°C to 25°C.
Salts used for desiccation and their corresponding mean water potentials are as follows: LiCl (-282 MPa), Mg(NO~3~)~2~ (-86 MPa), and NaCl (-38 MPa).
Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible; we therefore used the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described below.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Following the desiccation treatment, plants were rewetted using water from a local stream.
Plants were allowed to rehydrate, and chlorophyll fluorescence was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment chlorophyll fluorescence yield [@WatkinsJr.2007c].

@@

Rapid light response curves were constructed for each species by measuring photosynthetic yield at gradually increasing levels of photosynthetically active actinic light (400 nm to 700 nm) with the Light Curve function of the mini-PAM portable chlorophyll fluorometer (Walz Gmbh, Effeltrich, Germany) as described by the manufacturer.
Briefly, the maximum yield (F~m~) was first measured in the absence of actinic light.
Next, plants were equilibrated to the actinic light for 30 s, and photosynthetic yield (∆F/F~m~) measured with a 0.8 s saturating pulse.
This was repeated for each photosynthetic photon flux density (PPFD) level, up to ca.
500 µmol·m^-2^ ·s^-1^.
Response curves were fitted using the equation $y = A(1 - e^{-kx})$, where *y* is relative electron transport rate of photosystem II (RETR), *x* is PPFD, *A* is the asymptote of the curve, and *k* is a slope parameter; outliers at high PPFD likely to distort the curve were discarded [@Proctor2012].
The PPFD at 95% saturation of RETR (PPFD~95%~) and maximum RETR (ETR~max~) were then calculated from the curve.
Light responses were measured in 8--16 field-collected sporophyte individuals for each species, and on single gametophyte individuals in the field, which were later identified to species using the DNA barcodes.

@@

We used DNA barcoding to identify filmy fern gametophytes to species as described in @Nitta2017.
Briefly, DNA extraction was performed using a modified CTAB protocol [@Doyle1987] or the Plant Mini DNEasy kit (Qiagen, Valencia, California, USA) following the manufacturer's protocol.
Chloroplast *rbcL* and *trnH--psbA* were amplified and sequenced using primers and protocols following @Nitta2017.
For each marker, a reference library was constructed from sporophyte sequences using the makeblastDB command in BLAST [@Altschul1997].
Field-collected gametophytes were identified by a local BLAST query against the reference library.
Gametophytes matching >99.5% with a single reference sequence and no others were identified as that species; those that did not match any reference sequence or matched >99.5% with multiple reference sequences were not identified and excluded from further analysis.

@@

All statistical analyses were done in R v.4.0.2 [@RCoreTeam2020a].
Datasets and R scripts used in this study are available on Dryad (\#\#\#\#\#\#).

We derived the phylogenetic tree used in this study by extracting Hymenophyllaceae from the tree of @Nitta2017 with the extract.clade function of the ape package in R [@Paradis2004].
This tree is well-supported and generally in accord with previously published, more densely sampled phylogenies [@Pryer2001a; @Dubuisson2003; @Hennequin2006; @Ebihara2007a].

We quantified the degree of phylogenetic signal in physiological traits using Blomberg's *K* [@Blomberg2003] and Pagel's λ [@Pagel1999; @Freckleton2002] as implemented with the phylosig function in the phytools R package [@Revell2012].
Both measures test the hypothesis that the trait of interest is evolving according to Brownian Motion (BM).
λ is a scaling parameter that ranges from zero to one: λ near zero indicates random distribution of trait values across the tree, and λ near one indicates evolution of traits along the phylogeny following BM.
For *K*, values near one indicate evolution of traits following BM; *K* >1 indicates traits more conserved than expected under BM, and *K* <1 indicates that traits have less phylogenetic signal than expected under BM.
We tested the significance of *K* by comparing the observed value against values from a null distribution of 1000 phylogenies with the traits randomly shuffled across the tips.
We tested the significance of λ with a log-likelihood test comparing the likelihood of the observed value of λ vs. λ = 0.
For all further analyses, we used methods that account for phylogenetic history for physiological traits that showed significant phylogenetic signal, and standard methods otherwise.

We used general linear mixed models (GLMMs) to investigate the effect of growth habit and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at -86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (high-elevation epiphyte, low-elevation epiphyte, terrestrial or saxicolous) and generation (sporophyte vs. gametophyte) and their interactions as fixed effects.
We included species as a random effect in all models, and included a phylogenetic variance-covariance matrix in the case the response variable had significant phylogenetic signal (Hadfield, 2010).
We used an inverse-Gamma distribution for all priors, and ran analyses for 500,000 iterations, with burnin after 1,000 iterations and thinning every 500 iterations.
Comparisons of DT, ETR~max~, and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.

To test for a correlation between environmental water potential and DT, we first derived a single minimum water potential value for each species by taking the average of the minimum daily water potential at all sites where that species was present.
We used minimum values because these may represent a climatic extreme that limits the occurrence of species.
We used data from the 'epiphytic' dataloggers mounted at 1.5 m on trees for epiphytic species and data from the 'terrestrial' dataloggers on the ground for non-epiphytic species.
Data from the Mt. Rotui 830 m site were excluded because this site is much more exposed than the other high elevation sites, and appears as a significant outlier in the microclimatic data (see Results).
This may be due to its location; unlike other mountains on Moorea, Mt. Rotui is completely isolated with no connecting ridges to other peaks (Fig. 2).
Since DT showed significant phylogenetic signal (see Results), we used phylogenetic generalized least squares (PGLS) to test for a relationship between environmental water potential and DT while taking phylogeny into account, using species' means.

To test whether gametophytes with wider ranges have higher DT, we conducted PGLS with DT as the response variable and range width (measured as the difference between maximum and minimum observed elevation) as the explanatory variable. We ran PGLS using the caper R package [@Orme2018].

<!--

@@

--\>

## References

<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->

::: {#refs}
:::
